---
title: "Why fontalign"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Why fontalign}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = requireNamespace("ragg", quietly = TRUE) &&
         requireNamespace("svglite", quietly = TRUE)
)
```

`fontalign` exists because two problems show up together when rendering
plots on a Linux server for a global audience:

1. **Font size inconsistency across output formats.** The same
   `ggsave("plot.pdf", dpi = 300)` and `ggsave("plot.png", dpi = 300)`
   produce visibly different text heights when a font post-processor
   like **showtext** is enabled. This is because `showtext` calibrates
   its own internal DPI independently of the actual device DPI.

2. **Complex-script shaping.** Arabic, Persian, Devanagari, Thai, and
   emoji require HarfBuzz to render correctly. Cairo alone on a stock
   Linux image lacks the necessary Pango/Cairo bindings.

## What fontalign does and does not do

`fontalign`:

* Picks a backend (`cairo_pdf`, `cairo_png`, `ragg::agg_png`,
  `svglite::svglite`) appropriate for the scripts detected in the plot.
* Aligns text size to the target DPI across all chosen backends.
* Provides `fontalign_check_system()` so a system administrator can
  verify the server's font and graphics stack before deploying.

`fontalign` does **not**:

* Modify your ggplot object or theme.
* Globally `trace()` base graphics functions.
* Reimplement HarfBuzz shaping. That is delegated to `ragg`/`svglite`,
  which already do it correctly.

## Workflow

```{r, eval = FALSE}
library(fontalign)
fontalign_check_system()
```

```{r, eval = FALSE}
library(ggplot2)
p <- ggplot(mtcars, aes(mpg, wt)) +
     geom_point() +
     labs(title = "fuel consumption")

fontalign_save(p, "out.pdf", width = 6, height = 4, dpi = 300)
```

If the plot's labels contain Arabic or Thai, the same call routes to
`ragg::agg_png` for PNG output and `svglite::svglite` for SVG, so
shaping is handled by HarfBuzz via those backends.

## Server deployment

The recommended Linux baseline is:

```bash
apt-get install -y \
  libcairo2-dev libpango1.0-dev libfontconfig1-dev \
  libharfbuzz-dev libfribidi-dev librsvg2-dev \
  fonts-noto fonts-noto-cjk fonts-noto-color-emoji
```

After deploy, run `fontalign::fontalign_check_system()` from inside the
deployed environment to confirm.
